---
title: "Simple Models of Stock Return (R Version)"
output:
  html_document:
    df_print: paged
---

# **Models of Stock Return**

In this Notebook, I apply the concept of probability to measure the probability that the stock price drops a certain percentage in a day, and in a year.

I also demonstrate how to calculate the value at risk (VaR) using python.

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
```

```{r}
# Select CSV file interactively
data <- file.choose()
```

```{r}
fb <- read_csv(data, show_col_types = FALSE)
head(fb)
```

## **Distribution of Log return**

```{r}
# Calculate log daily return
fb <- fb %>%
  mutate(LogReturn = log(lead(Close)) - log(Close))
```

```{r}
# Estimate parameters
mu <- mean(fb$LogReturn, na.rm = TRUE)
sigma <- sd(fb$LogReturn, na.rm = TRUE)

# Create density data for the normal distribution
x_vals <- seq(min(fb$LogReturn, na.rm = TRUE) - 0.01,
              max(fb$LogReturn, na.rm = TRUE) + 0.01,
              by = 0.001)
pdf_vals <- dnorm(x_vals, mean = mu, sd = sigma)
density_df <- data.frame(x = x_vals, pdf = pdf_vals)

# Plot histogram and overlay normal distribution
ggplot(fb, aes(x = LogReturn)) +
  geom_histogram(aes(y = after_stat(density)), bins = 50, fill = "skyblue", color = "black", alpha = 0.7) +
  geom_line(data = density_df, aes(x = x, y = pdf), color = "red", size = 1) +
  labs(title = "Distribution of Facebook Log Returns",
       x = "Log Return", y = "Density") +
  theme_minimal()
```

## **Calculating the probability of the stock price dropping over a certain percentage in a day**

Use the python function: q = pnorm(z, mean = mu, sd = sigma).

**🎯 What the Function Tells You:**

"What’s the probability that a value from a normal distribution (with the given mean and standard deviation) is less than or equal to a certain point z?"

The result q is a number between 0 and 1 (a probability).

For example, if q = 0.11, it means:

“There’s an 11% chance the return will be -z% or worse.”

```{r}
# Probability that Facebook stock will drop more than 5% in a day
prob_return1 <- pnorm(-0.05, mean = mu, sd = sigma)
cat("The Probability is", prob_return1, "\n")
```

```{r}
# Calculate the probability that return drops over 10%
prob_return2 <- pnorm(-0.10, mean = mu, sd = sigma)
cat("The Probability is", prob_return2, "\n")
```

## **Calculating the probability of the stock price dropping over a certain percentage in a year**

```{r}
# Using `mu` and `sigma` defined above
mu220 <- 220 * mu
sigma220 <- sqrt(220) * sigma

# Compute probability of a drop > 40%
prob_drop <- pnorm(-0.4, mean = mu220, sd = sigma220)
cat("The probability of dropping over 40% in 220 days is", prob_drop, "\n")

```

```{r}
mu220 <- 220 * mu
sigma220 <- sqrt(220) * sigma

# Calculate probability of a drop over 20%
drop20 <- pnorm(-0.2, mean = mu220, sd = sigma220)

# Output the result
cat("The probability of dropping over 20% in 220 days is", drop20, "\n")
```

## **Calculating Value at Risk (VaR)**

Value at Risk (VaR) is a risk management metric used in finance to quantify the potential loss in value of a portfolio or investment under normal market conditions over a set time period and with a given level of confidence.

It answers the question:

> "What’s the maximum expected loss over a specific time horizon, with a certain level of confidence?"

**Limitations:**

1.  **Does not predict worst-case loss**, only a threshold.
2.  Assumes normal market conditions – **underestimates tail risk.**

A commonly used method to calculate VaR is the **Variance-Covariance (Parametric)**, which assumes returns are normally distributed (which is the case in my example above) and uses standard deviation.

Now let's estimate how much you could lose in a given time frame, **assuming returns are normally distributed.**

```{r}
# 5% quantile (VaR at 95% confidence level)
VaR <- qnorm(0.05, mean = mu, sd = sigma)
cat("Single day value at risk:", VaR, "\n")

# 25% quantile
q25 <- qnorm(0.25, mean = mu, sd = sigma)
cat("25% quantile:", q25, "\n")
```

VaR results on 5% quantile can be interpreted that there's a 5% chance that the return will be below 0.022523362407065 and a 95% chance it won't be worse than that. 😊
